---
title: "Mini Project 2"
author: "Evelyn Delph, Arthur Janes, Nick Mobley"
date: "3/26/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
# library imports go here
library(tidyverse)
library(GGally)
library(broom)
library(ggpubr)

# read in data
Source <- read.csv("DFP_WTHH_release.csv")
df = Source
colnames(df)[1] = "ID"
#Removed worthless columns
df = df %>%
  select(c(ID,presvote16post,house3,weight_DFP,M4A,
           GREENJOB,WEALTH,MARLEG,ICE,GUNS,
           POP_1,POP_2,POP_3))

df = df %>%
  mutate(
    Swinger = ifelse(presvote16post == house3 & 
                       presvote16post %in% c(1,2),FALSE,TRUE),
    Party = ifelse(presvote16post == house3 &
                     presvote16post %in% c(1,2),
                   ifelse(presvote16post == 1,"Democrat","Repulican"),
                   "Swinger")
  )

# remove rows with NA conflicts
df <- df[-c(2647, 3042, 3200),]

# separate data into groups based on voting pattern
loyalD = df %>%
  filter(presvote16post == 1 & house3 == 1)
loyalR = df %>%
  filter(presvote16post == 2 & house3 == 2) 

loyal <- rbind(loyalD, loyalR)

switchD = df %>%
  filter(presvote16post != 1 & house3 == 1) %>%
  mutate(Partyto = "Democrat")
switchR = df %>%
  filter(presvote16post != 2 & house3 == 2) %>%
  mutate(Partyto = "Republican")


switchers = rbind(switchD,switchR)



# subset data
issues = df %>%
  select(c(Swinger,M4A,GREENJOB,WEALTH,MARLEG,ICE,GUNS))
```

# How do Switch D and Switch R voters differ on issue variables?

```{r Plotting Function, include = F}
IssueSwitchPlot <- function(dataframe,issue) {
  ggplot(dataframe, aes(x = !!ensym(issue), 
                        group = Partyto, 
                        colour = Partyto,
                        fill = Partyto)) +
  geom_bar() + 
    facet_wrap(~Partyto,nrow = 2,scales = "free_y") +
    ylab("") + 
    xlab("") +
    scale_fill_manual(name = "Party Switched To",
                      values = c("#0015BC","#E9141D")) +
    scale_color_manual(name = "Party Switched To",
                       values = c("#0015BC","#E9141D"))  
}
```

```{r Setup Issue Plots,warning = FALSE,echo = FALSE,results= 'hide'}

wantedIssues = c("M4A","GREENJOB",'WEALTH','MARLEG','ICE','GUNS')



plot_list = lapply(wantedIssues, IssueSwitchPlot, dataframe = switchers)



figLabels = c("Medicare for All","Green Jobs Program","Wealth Tax",
              "Legalize Marijuana","Defunding ICE","Gun Control")


plotGrid = ggarrange(plotlist = plot_list,
          nrow = 3,
          ncol = 2,
          common.legend = TRUE,
          legend = "bottom",
          labels = figLabels)



```







```{r Plot Issues Plot,echo = FALSE, fig.width = 14, fig.height = 14,fig.align= 'center'}
annotate_figure(plotGrid,
                top = text_grob("How Swing Voters Feel on different Issues",
                                size = 36),
                bottom = text_grob("Survey Results on Scale 1-5, 1 = Strongly Support, 5 = Strongly Oppose (6 = \"Not Sure\")",size = 24))
```


#Section 2: How do Swing voters differ from loyal Republicans and loyal Democrats on issue variables?

```{r, warning = FALSE}


ggplot(df, aes(y = M4A, x = Party, color = Party)) + geom_boxplot() + coord_flip() + stat_summary(fun.y = mean, geom = "point", size = 4) + ggtitle("Medicare For ALL")

ggplot(df, aes(y = GREENJOB, x = Party, color = Party)) + geom_boxplot() + coord_flip() + stat_summary(fun.y = mean, geom = "point", size = 4) + ggtitle("A Green Jobs Program")

ggplot(df, aes(y = WEALTH, x = Party, color = Party)) + geom_boxplot() + coord_flip() + stat_summary(fun.y = mean, geom = "point", size = 4) + ggtitle("A tax on wealth over $100 million")

ggplot(df, aes(y = MARLEG, x = Party, color = Party)) + geom_boxplot() + coord_flip() + stat_summary(fun.y = mean, geom = "point", size = 4) + ggtitle("Legalizing Marijuana")

ggplot(df, aes(y = ICE, x = Party, color = Party)) + geom_boxplot() + coord_flip() + stat_summary(fun.y = mean, geom = "point", size = 4) + ggtitle("Defunding Immigration and Customs Enforcement")

ggplot(df, aes(y = GUNS, x = Party, color = Party)) + geom_boxplot() + coord_flip() + stat_summary(fun.y = mean, geom = "point", size = 4) + ggtitle("Gun Control")
```

#Section 3: What predicts being a swing voter?
There are two sets of variables that can determine swing voters from non-swing voters. First, the six issue variables will be examined, then the populism variables. After that, the best model among each variable group will be determined.

##Issue Variables
First, to get a glimspe at the data, a pairs plot was constructed to see how individual variables were related. Wealth and Medicare for All had the highest correlation. As a result, it was a assumed there was an interaction between them. The rest of the variables had similar correlations, but not quite as high as the interaction terms. Due to this, it was anticipated that all of the variables were important for predicting swing voters. 

For each model, the absolute difference between mean probabilities for swing voters and non swing voters were calculated. The model with the highest difference indicates a clearer separation between the two types of voters. THe model with the highest difference between the means is considered to be the best performing model as a result. Below shows the probability scores after adding each variable, as well as determining the most accurate model.
```{r,echo=F}

# adding variables to model at a time
one.issue <- glm(Swinger ~ WEALTH, family = quasibinomial, weights = weight_DFP, data = df)
two.issues <- glm(Swinger ~ WEALTH * M4A, family = quasibinomial, weights = weight_DFP, data = df)
three.issues <- glm(Swinger ~ WEALTH * M4A + MARLEG, family = quasibinomial, weights = weight_DFP, data = df)
four.issues <-  glm(Swinger ~ WEALTH * M4A + MARLEG + GUNS, family = quasibinomial, weights = weight_DFP, data = df)
five.issues <-  glm(Swinger ~ WEALTH * M4A + MARLEG + GUNS + GREENJOB, family = quasibinomial, weights = weight_DFP, data = df)
six.issues <- glm(Swinger ~ WEALTH * M4A + MARLEG + GUNS + GREENJOB + ICE, family = quasibinomial, weights = weight_DFP, data = df)

# probs for one issues
coe <- coef(one.issue)
y <- coe[1] + coe[2] * 1:5
probs1 <- exp(y) / (1 + exp(y)) 
cat("Probabilities for one issue variable (Wealth):\n",probs1)
print("")

# probs for two issues
coe <- coef(two.issues)
y <- coe[1] + coe[2] + coe[3] * coe[4] * 1:5
probs2 <- exp(y) / (1 + exp(y))
cat("\nProbabilities for two issue variables (WEALTH + M4A):\n",probs2)
print("")

coe <- coef(three.issues)
y <- coe[1] + coe[2] + coe[3] + coe[4] * coe[5] * 1:5
probs3 <- exp(y) / (1 + exp(y))
cat("\nProbabilities for three issue variables (WEALTH + M4A + MARLEG):\n",probs3)
print("")

coe <- coef(four.issues)
y <- coe[1] + coe[2] + coe[3] + coe[4] + coe[5] * coe[6] * 1:5
probs4 <- exp(y) / (1 + exp(y))
cat("\nProbabilities for four issue variables (WEALTH + M4A + MARLEG + GUNS):\n",probs4)
print("")

coe <- coef(five.issues)
y <- coe[1] + coe[2] + coe[3] + coe[4] + coe[5] + coe[6] * coe[7] * 1:5
probs5 <- exp(y) / (1 + exp(y))
cat("\nProbabilities for five issue variables (WEALTH + M4A + MARLEG + GUNS+ GREENJOB):\n",probs5)
print("")

coe <- coef(six.issues)
y <- coe[1] + coe[2] + coe[3] + coe[4] + coe[5] + coe[6] + coe[7] * coe[8] * 1:5
probs6 <- exp(y) / (1 + exp(y))
cat("\nProbabilities for six issue variables (WEALTH + M4A + MARLEG + GUNS+GREENJOB + ICE):\n",probs6)
```
```{r,echo=F}
swingers <- c(probs1,probs2,probs3,probs4,probs5,probs6)
best.model <- ""
last.dif <- 0
cat("\t\tNon-Swinger","Swinger","Difference","\n")

for(i in 1:6){
  not.swingers <- mean(1-swingers[i])
  swingers.mean  <- mean(swingers[i])
  dif <- abs(not.swingers-swingers.mean)
  
  if(dif > last.dif){
    last.df <- dif
    best.model <- i
  }
  
  cat("# Variables",i,not.swingers, swingers.mean, dif,"\n")
}

cat("\nThe best model that shows the greatest difference 
    between swing and non-swing votes the model with",i,"variables")
```
Final 6 variable equation:
```
votes = -1.99244285 + 0.29325492 + 0.18323905 + (-0.18073493) + 0.11441703  + 0.16567127 + (-0.07662718) * (-0.10216151 ) * X
```
For the issue variables, the model with all six variables seemed to have the highest separation with a difference of 0.6581869. This could show that all of the issues are important when determining a swing voter from a non-swing voter.

##Populism Variables
The same steps were conducted for the populism variables. The only difference was that there wasn't a significant interaction term among the variables. Below shows the probabilities and the selection for the most accurate model.

```{r,echo=F}
pop <- df[c("POP_1","POP_2","POP_3")]

# populism variable models
one.pop <- glm(Swinger ~ POP_1, family = quasibinomial, weights = weight_DFP, data = df)
two.pops <- glm(Swinger ~ POP_1 + POP_2, family = quasibinomial, weights = weight_DFP, data = df)
three.pops <- glm(Swinger ~ POP_1 + POP_2 + POP_3, family = quasibinomial, weights = weight_DFP, data = df)

# probs for one pop variable
coe <- coef(one.pop)
y <- coe[1] + coe[2] * 1:5
p1 <- exp(y) / (1 + exp(y)) 
cat("Probabilities for one populism variable (POP_1):\n",p1)
print("")

# probs for two pop variables
coe <- coef(two.pops)
y <- coe[1] + coe[2] + coe[3] * 1:5
p2 <- exp(y) / (1 + exp(y)) 
cat("Probabilities for two populism variables (POP_1+ POP_2):\n",p2)
print("")

# probs for three pop variables
coe <- coef(three.pops)
y <- coe[1] + coe[2] + coe[3] + coe[4] * 1:5
p3 <- exp(y) / (1 + exp(y)) 
cat("Probabilities for three populism variables (POP_1 + POP_2 + POP_3):\n",p3)
print("")
```
```{r,echo=F}
pops <- c(p1,p2,p3)
best.model.pop <- ""
last.dif.pop <- 0
cat("\t\tNon-Swinger","Swinger","Difference","\n")

for(i in 1:3){
  not.pop <- mean(1-pops[i])
  pop.mean  <- mean(pops[i])
  dif <- abs(not.pop-pop.mean)
  
  if(dif > last.dif.pop){
    last.df.pop <- dif
    best.model.pop <- i
  }
  
  
  cat("# Variables",i,not.pop,pop.mean,dif,"\n")
}

cat("\nThe best model that shows the greatest difference 
    between populism variables is the model with",i,"variables")

```
Final 3 populism variable equation:
```
votes = -1.06513692 + (-0.20446605) + (-0.04608700) + 0.06178131 * X
```
For the populism variables, the model with the three variables had the highest separation with a difference of 0.5555029. Again, this could show that encompassing all variables gives a holistic picture of the swing and non-swing voters' values.

##Best Model
Among the best two models, the issue variables out perform it by roughly 18%. This could mean that issue variables are slightly better predictors than populism variables, as they point to concrete issues rather than abstract beliefs and ideals.

#Appendix

##Section 3

GGPairs plot to detect correlation and possible interactions with issue variables:
```{r,echo=F}
ggpairs(issues)
```

Results of fitting the issue variable models:
```{r,echo=F}
# evaluate residuals
six.df <- augment(six.issues)

# evaluate final model
six.df$.fitted = fitted.values(six.issues)
six.df$.resid = residuals(six.issues, type="response") 

gg1 <- ggplot(six.df, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  geom_smooth(method = "loess", method.args = list(degree = 1)) + 
  xlab("Fitted values") + 
  ylab("Residuals") + 
  ggtitle("Fitted Values over Residuals")

gg2 <- ggplot(six.df, aes(x = WEALTH, y = .resid)) + 
  geom_point() + 
  geom_smooth(method = "loess", method.args = list(degree = 1)) + 
  xlab("Wealth") + 
  ylab("Residuals") + 
  ggtitle("Wealth Responses over Residuals")

gg3 <- ggplot(six.df, aes(x = M4A, y = .resid)) + 
  geom_point() + 
  geom_smooth(method = "loess", method.args = list(degree = 1)) + 
  xlab("M4A") + 
  ylab("Residuals") + 
  ggtitle("Medicare for All Responses over Residuals")

gg4 <- ggplot(six.df, aes(x = GUNS, y = .resid)) + 
  geom_point() + 
  geom_smooth(method = "loess", method.args = list(degree = 1)) + 
  xlab("Guns") + 
  ylab("Residuals") + 
  ggtitle("Gun Responses over Residuals")

gg5 <- ggplot(six.df, aes(x = ICE, y = .resid)) + 
  geom_point() + 
  geom_smooth(method = "loess", method.args = list(degree = 1)) + 
  xlab("ICE") + 
  ylab("Residuals") + 
  ggtitle("ICE Responses over Residuals")

gg6 <- ggplot(six.df, aes(x = GREENJOB, y = .resid)) + 
  geom_point() + 
  geom_smooth(method = "loess", method.args = list(degree = 1)) + 
  xlab("GREENJOB") + 
  ylab("Residuals") + 
  ggtitle("Green-job Responses over Residuals")
  

gg7 <- ggplot(six.df, aes(x = MARLEG, y = .resid)) + 
  geom_point() + 
  geom_smooth(method = "loess", method.args = list(degree = 1)) + 
  xlab("MARLEG") + 
  ylab("Residuals") +
  ggtitle("Legalizing Marijuana Responses over Residuals")

gg1
gg2
gg3
gg4
gg5
gg6
gg7
```

GGPairs for populism variables:
```{r,echo=F}
ggpairs(pop)
```

Results from fitting populism variables:
```{r,echo=F}
# evaluate residuals
pop.df <- augment(three.pops)

# evaluate final model
pop.df$.fitted = fitted.values(three.pops)
pop.df$.resid = residuals(three.pops, type="response") 

gg1 <- ggplot(pop.df, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  geom_smooth(method = "loess", method.args = list(degree = 1)) + 
  xlab("Fitted values") + 
  ylab("Residuals") + 
  ggtitle("Fitted Values over Residuals")

gg2 <- ggplot(pop.df, aes(x = POP_1, y = .resid)) + 
  geom_point() + 
  geom_smooth(method = "loess", method.args = list(degree = 1)) + 
  xlab("POP_1") + 
  ylab("Residuals") + 
  ggtitle("POP 1 Variable over Residuals")

gg3 <- ggplot(pop.df, aes(x = POP_2, y = .resid)) + 
  geom_point() + 
  geom_smooth(method = "loess", method.args = list(degree = 1)) + 
  xlab("POP_2") + 
  ylab("Residuals") + 
  ggtitle("POP 2 Variable over Residuals")

gg4 <- ggplot(pop.df, aes(x = POP_3, y = .resid)) + 
  geom_point() + 
  geom_smooth(method = "loess", method.args = list(degree = 1)) + 
  xlab("POP_3") + 
  ylab("Residuals") + 
  ggtitle("POP 3 Variable over Residuals")


gg1
gg2
gg3
gg4
```



